\documentclass[twoside,leqno,twocolumn]{article}

\usepackage[letterpaper]{geometry}
\usepackage{ltexpprt}
%\usepackage[max6]{authblk}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{algpseudocode}
\usepackage{capt-of}
%\usepackage{ntheorem}
% for appendix
\usepackage{xspace}
\usepackage{mathtools}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\newtheorem{definition}{Definition}
\newtheorem{claim}{Claim}
\newtheorem{myfunc}{{Functionality}}
\newtheorem{myalgorithm}{{Algorithm}}
\renewcommand{\paragraph}[1]{\medskip\noindent\textbf{#1}}

\newcommand{\X}{\mathbf{X}}
\renewcommand{\sec}{{\lambda}}
\newcommand{\e}{{\epsilon}}
\newcommand{\A}{{\cal A}}
\newcommand{\bsize}{\ensuremath{b}}
\newcommand{\op}{\ensuremath{\mathsf{op}}\xspace}
\newcommand{\data}{{\ensuremath{\mathsf{data}}}\xspace}
\newcommand{\addr}{{\ensuremath{\mathsf{addr}}}\xspace}
\newcommand{\bit}{{\{0,1\}}}
\newcommand{\Sim}{\ensuremath{{\sf Sim}}\xspace}


\newenvironment{MyItemize}[1]{\begin{itemize}\setlength{\itemsep}{0.1cm}
\setlength{\parskip}{-0.05cm} 
\setlength{\leftmargin}{5mm}
#1}{\end{itemize}}




\newenvironment{MyEnumerate}[1]{
\begin{enumerate}
\setlength{\itemsep}{0.1cm}
\setlength{\parskip}{-0.05cm} 
\setlength{\leftmargin}{5mm}
#1}{\end{enumerate}}

\usepackage{color,xcolor}
\usepackage{flushend}

\ifpdf
\usepackage[bookmarks=true,pdfstartview=FitH,colorlinks,linkcolor=black,filecolor=black,citecolor=black,urlcolor=black]{hyperref}

\else

\usepackage[dvipdfm,bookmarks=true,pdfstartview=FitH,colorlinks,linkcolor=black,filecolor=black,citecolor=black,urlcolor=black]{hyperref}

\fi


\input{macro}


\sloppy
\widowpenalty10000
\clubpenalty10000

\begin{document}

\title{\bf Bucket Oblivious Sort: \\An Extremely Simple Oblivious Sort}

\author{Gilad Asharov \thanks{Bar-Ilan University. Part of the work was done while the author was a post-doctoral fellow at Cornell Tech supported a Junior Fellow award from the Simons Foundation, and while at J.P. Morgan AI Research.} \quad
T-H. Hubert Chan\thanks{The University of Hong Kong. Partially supported the Hong Kong RGC under the grant 17200418.} \quad
Kartik Nayak\thanks{Duke University. Part of the work was done while the author was at University of Maryland.} \quad
Rafael Pass\thanks{Cornell Tech.} \quad
Ling Ren\thanks{University of Illinois Urbana-Champaign. Part of the work was done while the author was at MIT.} \quad
Elaine Shi\thanks{Cornell University.}}

\newcommand{\rl}[1]{{\footnotesize\color{orange}[Ling: #1]}}

\date{}

\maketitle

\fancyfoot[R]{\scriptsize{Copyright \textcopyright\ 2020 by SIAM\\
Unauthorized reproduction of this article is prohibited}}

\begin{abstract}
\input{abstract}
\end{abstract}

\input{intro}
\input{defn}
\input{algo}
\input{extension}

\bibliographystyle{plain}
\bibliography{refs}


\appendix

%\input{locality}

\end{document}
